Supplemental material: Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-03-15 Manuscript/HI_Model_ms.Rmd
updated manuscript and supplementary files

### *4.3 Limitations of this study* 
There is ample evidence that increasing atmospheric CO~2~ concentrations can stimulate NPP [@norby_forest_2005; @zhu_greening_2016], but the magnitude and persistence of this effect remains highly uncertain, especially over long time spans [@franks_sensitivity_2013; @walker_integrating_2020]. Our results demonstrate that long-term projections of ecosystem carbon balance are highly sensitive to uncertainty in CFE strength. With no CFE, Hawaiian ecosystems transitioned into a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario. However, a CFE equivalent to a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems remaining a net carbon sink throughout the entire simulation period. A 15% CFE applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations. Despite this demonstrated sensitivity to a CFE, the combined effect of several potentially attenuating factors makes it difficult to determine the appropriate magnitude of a CFE with any degree of confidence [@walker_integrating_2020]. Nitrogen and phosphorus limitation can reduce or eliminate a CFE [@reich_nitrogen_2006; @norby_co2_2010; @he_nitrogen_2017; @terrer_nitrogen_2019], as can water limitation and heat stress [@obermeier_reduced_2017; @birami_hot_2020]. Forest age may also be a factor, with young aggrading forests showing a strong positive growth response to CO~2~ fertilization [@walker_decadal_2019], while old-growth forests show little to no response [@jiang_fate_2020; @yang_low_2020]. This evidence indicates that a CFE may be highly variable across space and time, suggesting that it may be unrealistic to apply a single CFE rate value across an entire region over several decades. Until mechanistic understanding is improved, the most conservative approach when projecting future ecosystem balance in the context of climate mitigation planning may be to assume no CFE, with the knowledge that any realized CFE will only enhance ecosystem carbon sequestration.

>>>>

### *4.3 Limitations of this study* 
There is ample evidence that increasing atmospheric CO~2~ concentrations can stimulate NPP [@norby_forest_2005; @zhu_greening_2016], but the magnitude and persistence of this effect remains highly uncertain [@franks_sensitivity_2013; @walker_integrating_2020]. Our results demonstrate that long-term projections of ecosystem carbon balance are highly sensitive to uncertainty in CFE strength. With no CFE, Hawaiian ecosystems transitioned into a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario. However, a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems remaining a net carbon sink throughout the entire simulation period. A 15% CFE applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations. Despite this demonstrated sensitivity to a CFE, the combined effect of several potentially attenuating factors makes it difficult to determine the appropriate magnitude of a CFE with any degree of confidence [@walker_integrating_2020]. Nitrogen and phosphorus limitation can reduce or eliminate a CFE [@reich_nitrogen_2006; @norby_co2_2010; @he_nitrogen_2017; @terrer_nitrogen_2019], as can water limitation and heat stress [@obermeier_reduced_2017; @birami_hot_2020]. Forest age may also be a factor, with young aggrading forests showing a strong positive growth response to CO~2~ fertilization [@walker_decadal_2019], while old-growth forests show little to no response [@jiang_fate_2020; @yang_low_2020]. This evidence indicates that a CFE may be highly variable across space and time, suggesting that it may be unrealistic to apply a single CFE rate value across an entire region over several decades. Until mechanistic understanding is improved, the most conservative approach when projecting future ecosystem balance in the context of climate mitigation planning may be to assume no CFE, with the knowledge that any realized CFE will only enhance ecosystem carbon sequestration.

### 2021-03-15 Manuscript/HI_Model_supp.Rmd
updated manuscript and supplementary files

```{r fig.align = 'center', out.width = "99%", fig.cap = "Annual area burned by wildland fire in the State of Hawaiʻi from 1999-2019, summed by year across the seven main Hawaiian Islands (A) and within each of the four largest islands (B). The dashed horizontal line in (A) represents the median area burned from 1999-2019."}
knitr::include_graphics("./fig_images/figS2_AnnualAreaBurned.png")
```
## Climate

Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014. Spatially explicit projections of changes in annual temperature and rainfall by the year 2100 were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCPs 4.5 and 8.5 (Figure S5).

@@

Projections of total land area covered by each State Class were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands from LUCAS model output data.

```{r fig.align = 'center', out.width = "99%", fig.cap = "Projections of total land area by year for each of five State Classes in the seven main Hawaiian Islands (Agriculture, Developed, Forest, Grassland, and Shrubland) under low and high land use change scenarios for the period 2010-2100. Solid lines represent the mean of 30 Monte Carlo realizations and shaded areas represent minimum and maximum Monte Carlo values."}
knitr::include_graphics("./fig_images/figS5_StateClassArea.png")
```
## References

>>>>

```{r fig.align = 'center', out.width = "99%", fig.cap = "Annual area burned by wildland fire in the State of Hawaiʻi from 1999-2019, summed by year across the seven main Hawaiian Islands (A) and within each of the four largest islands (B). The dashed horizontal line in (A) represents the median area burned from 1999-2019."}
knitr::include_graphics("./fig_images/figS2_AnnualAreaBurned.png")
```

## Climate

Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014. Spatially explicit projections of changes in annual temperature and rainfall by the year 2100 (Figure S4) were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCPs 4.5 and 8.5.

@@

Projections of total land area covered by each of five State Classes (Agriculture, Developed, Forest, Grassland and Shrubland) were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios  (low and high; Figure S5).

```{r fig.align = 'center', out.width = "99%", fig.cap = "Projections of total land area by year for each of five State Classes in the seven main Hawaiian Islands (Agriculture, Developed, Forest, Grassland, and Shrubland) under low and high land use change scenarios for the period 2010-2100. Solid lines represent the mean of 30 Monte Carlo realizations and shaded areas represent minimum and maximum Monte Carlo values."}
knitr::include_graphics("./fig_images/figS5_StateClassArea.png")
```

## References
